Supporting Information to Estimating the Initial Susceptible Fraction in Dengue Dynamics"

### 2014-11-27 S1_lm1.tex
continuing the description of the stochastic model

In order to validate the estimates derived in the paper, we simulated 
multi-strain Dengue dynamics by means of a stochastic 4-serotype SIR model with 
cross-immunity as described below. The model permits up to 4 dengue infections 
with reduced susceptibility after the first Dengue episode due to cross 
immunity. Immunity to each serotype is considered complete and permanent. 
Figure xx depicted all possible states and state-transitions included in the 
model.

@@

The possible state-transitions and their  propensities are listed in table 
\ref{tab:trans}.
\begin{table}
\caption{
\bf{State-transitions and propensities}}
\label{tab:trans}
\begin{center}
\begin{tabular}[c]{l|l|l}
\hline
Transition & Propensity & Description\\
\hline
$S \rightarrow I_i$ & $\beta S I_{*i}$ & Primary infection \\
$I_i \rightarrow R_i$ & $\sigma I_i$ & Primary recovery\\
$R_i \rightarrow I_{[i]j}$ & $\beta \delta R_i I_{*j}$ & Secondary infection\\
$I_{[i]j} \rightarrow R_{ij}$ & $\sigma I_{[i]j}$ & Secondary recovery\\
$R_{ij} \rightarrow I_{[ij]k}$ & $\beta \delta R_{ij} I_{*k}$ & Tertiary 
infection\\
$I_{[ij]k} \rightarrow R_{ijk}$ & $\sigma I_{[ij]k}$ & Tertiary recovery\\
$R_{ijk} \rightarrow I_{[ijk]l}$ & $\beta \delta R_{ijk} I_{*l}$ & Quaternary 
infection\\
$I_{[ijk]l} \rightarrow R_{ijkl}$ & $\sigma I_{[ijk]l}$ & Quaternary recovery\\
$\rightarrow S$ & $\mu N$ & Birth\\
$All \rightarrow$ & $\mu N$ & Death\\
\hline
\end{tabular}
\end{center}

@@

>>>>

In order to validate the estimates derived in the paper, we simulated 
multi-strain Dengue dynamics by means of a stochastic 4-serotype SIR model with 
cross-immunity as described below. 
The model permits up to 4 dengue infections 
with reduced susceptibility after the first Dengue episode due to cross 
immunity. 
Immunity to each serotype is considered complete and permanent. 
Figure xx depicts all possible states and state-transitions included in the 
model.

          \begin{figure}
 \centering
 \includegraphics[width=16cm]{Dengue4.png}

 \caption{Block diagram detailing the stochastic model}
 \label{fig:sde_blocks}
\end{figure}

@@

The possible state-transitions and their  propensities are listed in table 
\ref{tab:trans}.
\begin{table}
\caption{
\bf{State-transitions and propensities: $P(\Delta X(t)|X(t))$}. The 
transitions are summarized below, Fully expanded the system comtemplates 64 
possible state transitions, as can be verified in figure \ref{fig:sde_blocks}. 
}
\label{tab:trans}
\begin{center}
\begin{tabular}[c]{l|l|l|l}
\hline
Transition & Propensity & State Change & Description\\
\hline
$S \rightarrow I_i$ & $\beta S I_{*i}$ & $\Delta S(t)=-1,\, \Delta I_i(t) = 1$ 
& Primary infection \\
$I_i \rightarrow R_i$ & $\sigma I_i$ & $\Delta I_i(t)=-1,\, \Delta R_i(t) = 1$ 
& Primary recovery\\
$R_i \rightarrow I_{[i]j}$ & $\beta \delta R_i I_{*j}$ & $$\Delta R_i(t)=-1,\, 
\Delta I_{[i]j}(t) = 1$$ & Secondary infection\\
$I_{[i]j} \rightarrow R_{ij}$ & $\sigma I_{[i]j}$ &$$\Delta I_{[i]j}(t)=-1,\, 
\Delta R_{ij}(t) = 1$$& Secondary recovery\\
$R_{ij} \rightarrow I_{[ij]k}$ & $\beta \delta R_{ij} I_{*k}$ & Tertiary 
infection\\
$I_{[ij]k} \rightarrow R_{ijk}$ & $\sigma I_{[ij]k}$ & Tertiary recovery\\
$R_{ijk} \rightarrow I_{[ijk]l}$ & $\beta \delta R_{ijk} I_{*l}$ & Quaternary 
infection\\
$I_{[ijk]l} \rightarrow R_{ijkl}$ & $\sigma I_{[ijk]l}$ & Quaternary recovery\\
$\rightarrow S$ & $\mu N$ & Birth\\
$All \rightarrow$ & $\mu N$ & Death\\
\hline
\end{tabular}
\end{center}

@@

The model is implemented as a continuous time Markov jump process. Let
$$\overrightarrow{X}(t) = [S(t), I_1(t), I_2(t), \ldots, R_{1234}(t)]$$ 
be the 
state of the system at the time $t$. 
The system is written as a forward Kolmogorov differential equation, which in 
matrix form looks like
\begin{equation}
dP_X(t) = Q P_X(t) 
\end{equation}

Where $P_X(t)$ is  the matrix of transition probabilities (given in table 
\ref{tab:trans}) and Q is the generator matrix, whose non-zero values are also 
given in table \ref{tab:trans} (in the state change column). The full formula 
and matrices are ommited due to their large sizes.
